---
title: "Prediction of ozone level in Boston"
author: 
- Lucas Emanuel Resck Domingues
- Lucas Machado Moscheb^[Escola de Matemática Aplicada]
output:
  pdf_document: default
  html_notebook: default
---

# Predicting O3 in Boston

```{r echo = F, warning = F, message=F}
library(zoo)
library(dplyr)
library(forecast)
library(tseries)
```


## Load and visualize

```{r, echo = F}
bos = read.csv("data/BOS.csv")
plot(bos$Date.Local, bos$O3.Mean, main = 'Daily average level of O3 in Boston', 
                                  xlab = 'Date', ylab = 'Kg/m3')
```

## Data treatment

We noticed that some days do not exist in the dataset, for example, the day August 31, 2001 does not have information in the dataset. 

```{r, echo = F}
bos[148:153,]
```

Also, there is duplicated days, as June 9, 2002: 

```{r, echo=F}
day <- as.Date(bos$Date.Local[duplicated(bos$Date.Local)])
bos[bos$Date.Local %in% c(toString(day - 1), toString(day), toString(day + 1)),]
```

The duplicated one is easier to deal, but the nan values are harder. First we calculate the mean value between the duplicated. 

```{r, echo = F}
bos[413, "O3.Mean"] = mean(bos[bos$Date.Local == toString(day),]$O3.Mean)
bos <- distinct(bos, Date.Local, .keep_all = TRUE)
```

We use `zoo` and `tseries` libraries in order to handle the data from now on. We also separate the train and test (last 365 days).

```{r, echo = F}
o3 <- zooreg(bos$O3.Mean, order.by = as.Date(as.character(bos$Date.Local), format = "%Y-%m-%d"))
o3.ts <- as.ts(o3)
```

```{r, echo= F}
o3_train = o3[1:(length(o3)[1] - 365),]
o3_test = o3[(length(o3)[1] - 365 + 1):length(o3)[1],]
```

## Models

Now we develop some models using the train data.

The metric to compare is the Mean Absolute Error (MAE) in the predictions: 

```{r, echo=F}
mae <- function(error)
{
    mean(abs(error))
}
```

### Decompose

```{r, echo=F}
# STL work only with tseries object 
#loess_model <- slt()
```


### Regression

### Holt-Winters

### ARMA

